---
title: "Homework 4"
author: "Zhan, Zhuoyi"
date: "`r date()`"
output: github_document
---

 Question 4.4

 (a) On average, what fraction of the available observations will we use to make the prediction?

 Becuase there is only 1 feature and it is uniformly distributed, so on average we will use 10% of the obervation


 (b)

 1%


 (c)

0.01^100 *100 = 10^-98 %


(d)

The fraction of training data near test test observation decrease exponentially.

(e)
 l = 0.1^(1/p)
the volumn of hypercube’ will be 10% of the total volume of space in the p-dimension cube.


Question 4.10

a)
```{r}
#install.packages("ISLR")
#install.packages('caret')
library(ISLR)
library(caret)
library(MASS)
library(class)
library(e1071)
summary(Weekly)
```

```{r}
pairs(Weekly)
```
Dimension and year may have a curved, non-linear pattern. and direction has linear pattern with other vriables.

b)
```{r}
lr.fit = glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Weekly, 
    family = binomial)
summary(lr.fit)
```
Lag 2 has a pr(>|z|) of 0.0296 which is statistically significant.

c)
```{r}
glm.probs = predict(lr.fit, type = "response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs > 0.5] = "Up"
table(glm.pred, Weekly$Direction)
```
The  Accuracy rate is (54+557)/(54+557+48+430) = 56.1%.
The True positive rate is 557/(48+557) = 92.1% So 92.1% of time, the logistic regression predict as up is right.

d)
```{r}
train = (Weekly$Year < 2009)
hold = Weekly[!train, ]
glm.fit = glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)
glm.probs = predict(glm.fit, hold, type = "response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs > 0.5] = "Up"
Direction.hold = Weekly$Direction[!train]
table(glm.pred, Direction.hold)
mean(lda.pred$class == Direction.hold)
```

overall fraction of correct predictions is (56+9)/(5+56+34+9) = 62.5%

e)
```{r}
lda.fit = lda(Direction ~ Lag2, data = Weekly, subset = train)
lda.pred = predict(lda.fit, hold)
table(lda.pred$class, Direction.hold)
mean(lda.pred$class == Direction.hold)
```

overall fraction of correct predictions is 62.5%

f)
```{r}
qda.fit = qda(Direction ~ Lag2, data = Weekly, subset = train)
qda.class = predict(qda.fit, hold)$class
table(qda.class, Direction.hold)
mean(qda.class == Direction.hold)

```
overall fraction of correct predictions is 58.65%

g)
```{r}
train.X = as.matrix(Weekly$`Lag2`[train])
test.X = as.matrix(Weekly$`Lag2`[!train])
train.Direction = Weekly$Direction[train]
set.seed(1)
knn.pred = knn(train.X, test.X, train.Direction, k = 1)
table(knn.pred, Direction.hold)
mean(knn.pred == Direction.hold)
```
overall fraction of correct predictions is 50%

h)
```{r}
nb.fit = naiveBayes(Direction ~ Lag2, data = Weekly, subset = train)
nb.class = predict(nb.fit, hold)
table(nb.class, Direction.hold)
mean(nb.class == Direction.hold)
```

i)
```{r}
# LOgistic regression and LDA 
```

j)
```{r}
# KNN k =10
knn.pred = knn(train.X, test.X, train.Direction, k = 10)
table(knn.pred, Direction.hold)
mean(knn.pred == Direction.hold)
```
```{r}
# KNN k =50
knn.pred = knn(train.X, test.X, train.Direction, k = 50)
table(knn.pred, Direction.hold)
mean(knn.pred == Direction.hold)
```

```{r}
#Logistic regression with sqrt(abs(Lag2))

glm.fit = glm(Direction ~ Lag2 + sqrt(abs(Lag2)), data = Weekly, family = binomial, subset = train)
glm.probs = predict(glm.fit, hold, type = "response")
glm.pred = rep("Down", length(glm.probs))
glm.pred[glm.probs > 0.5] = "Up"
Direction.hold = Weekly$Direction[!train]
table(glm.pred, Direction.hold)
mean(glm.pred == Direction.hold)
```
```{r}
# LDA with Lag2 interaction with Lag1
lda.fit = lda(Direction ~ Lag2:Lag1, data = Weekly, subset = train)
lda.pred = predict(lda.fit, hold)
table(lda.pred$class, Direction.hold)
mean(lda.pred$class == Direction.hold)
```

