---
title: "Homework 1"
author: "Zhan, Zhuoyi"
date: "`r date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('class')
library('dplyr')

## load binary classification example data from author website 
## 'ElemStatLearn' package no longer available
load(url('https://web.stanford.edu/~hastie/ElemStatLearn/datasets/ESL.mixture.rda'))
dat <- ESL.mixture

plot_mix_data <- expression({
  plot(dat$x[,1], dat$x[,2],
       col=ifelse(dat$y==0, 'blue', 'orange'),
       pch=20,
       xlab=expression(x[1]),
       ylab=expression(x[2]))
  ## draw Bayes (True) classification boundary
  prob <- matrix(dat$prob, length(dat$px1), length(dat$px2))
  cont <- contourLines(dat$px1, dat$px2, prob, levels=0.5)
  rslt <- sapply(cont, lines, col='purple')
})

eval(plot_mix_data)
```

```{r}
## fit linear classifier
fit_lc <- function(y, x) {
  x <- cbind(1, x)
  beta <- drop(solve(t(x)%*%x)%*%t(x)%*%y)
}

## make predictions from linear classifier
predict_lc <- function(x, beta) {
  cbind(1, x) %*% beta
}

## fit model to mixture data and make predictions
lc_beta <- fit_lc(dat$y, dat$x)
lc_pred <- predict_lc(dat$xnew, lc_beta)

## reshape predictions as a matrix
lc_pred <- matrix(lc_pred, length(dat$px1), length(dat$px2))
contour(lc_pred,
      xlab=expression(x[1]),
      ylab=expression(x[2]))


## find the contours in 2D space such that lc_pred == 0.5
lc_cont <- contourLines(dat$px1, dat$px2, lc_pred, levels=0.5)

## plot data and decision surface
eval(plot_mix_data)
sapply(lc_cont, lines)

```

Homework 1
Using the RMarkdown/knitr/github mechanism, implement the following tasks by extending the example R script mixture-data-lin-knn.R:
Paste the code from the mixture-data-lin-knn.R file into the homework template Knitr document.
Read the help file for R's built-in linear regression function lm
Re-write the functions fit_lc and predict_lc using lm, and the associated predict method for lm objects.
Consider making the linear classifier more flexible, by adding squared terms for x1 and x2 to the linear model
Describe how this more flexible model affects the bias-variance tradeoff



# Question 1
```{r}
data <- data.frame(dat$y,dat$x)
beta_lm <- lm(dat.y~., data)
summary(beta_lm)
#newx <-data.frame(dat$xnew)

```

```{r}
newx <-data.frame(dat$xnew)
newx <- rename(newx, X1= x1, X2=x2)
pred <- predict.lm(beta_lm, newx)
```

# Question 2

```{r}
sq_data <- data.frame(dat$y,dat$x,(dat$x)**2)
sq_beta_lm <- lm(dat.y~., sq_data)
sq_x <-data.frame(dat$xnew,(dat$xnew)**2)
sq_x <- rename(sq_x, X1= x1, X2=x2,X1.1=x1.1, X2.1=x2.1)
sq_pred <- predict.lm(sq_beta_lm, sq_x)
summary(sq_beta_lm)


```
# Question 3

Error is the combination of bias and variance. As the model gets more flexible, we are reducing testing error which means the model is fitted better, But when the level of complexity gets higher and higher, the model will overfit and the test error goes higher again. So the bias-variance tradeoff means that we need to reach the optimal level of flexibility, to minimize the test error.






```{r}

## fit knn classifier
## use 5-NN to estimate probability of class assignment
knn_fit <- knn(train=dat$x, test=dat$xnew, cl=dat$y, k=5, prob=TRUE)
knn_pred <- attr(knn_fit, 'prob')
knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)

## reshape predictions as a matrix
knn_pred <- matrix(knn_pred, length(dat$px1), length(dat$px2))
contour(knn_pred,
        xlab=expression(x[1]),
        ylab=expression(x[2]),
        levels=c(0.2, 0.5, 0.8))


## find the contours in 2D space such that knn_pred == 0.5
knn_cont <- contourLines(dat$px1, dat$px2, knn_pred, levels=0.5)

## plot data and decision surface
eval(plot_mix_data)
sapply(knn_cont, lines)

```

```{r}

## do bootstrap to get a sense of variance in decision surface
resample <- function(dat) {
  idx <- sample(1:length(dat$y), replace = T)
  dat$y <- dat$y[idx]
  dat$x <- dat$x[idx,]
  return(dat)
}
  
## plot linear classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  ## fit model to mixture data and make predictions
  lc_beta <- fit_lc(datb$y, datb$x)
  lc_pred <- predict_lc(datb$xnew, lc_beta)
  
  ## reshape predictions as a matrix
  lc_pred <- matrix(lc_pred, length(datb$px1), length(datb$px2))

  ## find the contours in 2D space such that lc_pred == 0.5
  lc_cont <- contourLines(datb$px1, datb$px2, lc_pred, levels=0.5)
  
  ## plot data and decision surface
  eval(plot_mix_data)
  sapply(lc_cont, lines)
}

```

```{r}

## plot 5-NN classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  
  knn_fit <- knn(train=datb$x, test=datb$xnew, cl=datb$y, k=5, prob=TRUE)
  knn_pred <- attr(knn_fit, 'prob')
  knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)
  
  ## reshape predictions as a matrix
  knn_pred <- matrix(knn_pred, length(datb$px1), length(datb$px2))

  ## find the contours in 2D space such that knn_pred == 0.5
  knn_cont <- contourLines(datb$px1, datb$px2, knn_pred, levels=0.5)
  
  ## plot data and decision surface
  eval(plot_mix_data)
  sapply(knn_cont, lines)
}

```

```{r}

## plot 20-NN classifier for three bootstraps
par(mfrow=c(1,3))
for(b in 1:3) {
  datb <- resample(dat)
  
  knn_fit <- knn(train=datb$x, test=datb$xnew, cl=datb$y, k=20, prob=TRUE)
  knn_pred <- attr(knn_fit, 'prob')
  knn_pred <- ifelse(knn_fit == 1, knn_pred, 1-knn_pred)
  
  ## reshape predictions as a matrix
  knn_pred <- matrix(knn_pred, length(datb$px1), length(datb$px2))
  
  ## find the contours in 2D space such that knn_pred == 0.5
  knn_cont <- contourLines(datb$px1, datb$px2, knn_pred, levels=0.5)
  
  ## plot data and decision surface
  eval(plot_mix_data)
  sapply(knn_cont, lines)
}
```

