---
title: "Homework 3"
author: "Zhan, Zhuoyi"
date: "`r date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Use the prostate cancer data. Use the cor function to reproduce the correlations

```{r}
## load prostate data
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))

cor(prostate[,1:8], use= "pairwise.complete.obs")
```

```{r}
#prostate_train <- subset(prostate, train==TRUE) %>% select(-'train')
#test <- subset(prostate, train==FALSE) %>% select(-'train')
```

### Treat lcavol as the outcome, and use all other variables in the data set as predictors.
```{r}
library(tidyverse)
library('splines')        ## for 'bs'
library('dplyr')  
library('magrittr')
library(glmnet)    
proTrain <- filter(prostate, train)
proTest <- filter(prostate, !train)
proTrain <-select(proTrain, -train)
proTest <-select(proTest, -train)
```

### With the training subset of the prostate data, train a least-squares regression model with all predictors using the lm function.

```{r}

fit <- lm(lcavol ~ ., 
             data=proTrain)
summary(fit)
```


```{r}
#prostate.lm <- lm(lpsa ~ ., data=prostate_train)
#summary(prostate.lm)
```

### Use the testing subset to compute the test error (average squared-error loss) using the fitted least-squares regression model.

```{r}
predictions <- predict(fit, newdata=proTest)
((predictions  -  proTest$lcavol)^2 %>% sum)/ length(predictions) 
```


###Train a ridge regression model using the glmnet function, and tune the value of lambda (i.e., use guess and check to find the value of lambda that approximately minimizes the test error).



```{r}
L2_loss <- function(y, yhat)
  (y-yhat)^2
error <- function(dat, fit, loss=L2_loss)
  mean(loss(dat$lcavol, predict(fit, newdata=dat)))

x_inp <- model.matrix(lcavol ~  lweight + age + lbph + lcp + pgg45 + lpsa + svi + gleason, data=proTrain)
y_out <- proTrain$lcavol
fit <- glmnet(x=x_inp, y=y_out, alpha=0, lambda=seq(0.5, 0, -0.05))
print(fit$beta)
```

```{r}
form  <- lcavol ~  lweight + age + lbph + lcp + pgg45 + lpsa + svi + gleason
error <- function(dat, fit, lam, form, loss=L2_loss) {
  x_inp <- model.matrix(form, data=dat)
  y_out <- dat$lcavol
  y_hat <- predict(fit, newx=x_inp, s=lam)  ## see predict.elnet
  mean(loss(y_out, y_hat))
}
```



```{r}
library(dplyr)
library(tidyverse)
min_error <- Inf
count <- 0
for (lam in seq(0,1,0.01)){
  err <-  error(proTest, fit, lam, form=form)
  count <- count + 1
  if (err < min_error){
    min_error = err
    cur_lambda = lam
  }
}
(cur_lambda)

#error(proTest, fit, lam=0, form=form)
#error(proTest, fit, lam=0.03, form=form)
#error(proTest, fit, lam=0.05, form=form)
#error(proTest, fit, lam=0.1, form=form)
#error(proTest, fit, lam=0.12, form=form)
```

###  Create a figure that shows the training and test error associated with ridge regression as a function of lambda. Create a path diagram of the ridge regression analysis, similar to HTF Figure 3.8

```{r}
err_train_1 <- sapply(fit$lambda, function(lam) 
  error(proTrain, fit, lam, form))
err_test_1 <- sapply(fit$lambda, function(lam) 
  error(proTest, fit, lam, form))
```

```{r}
plot(x=range(fit$lambda),
     y=range(c(err_train_1, err_test_1)),
     xlim=rev(range(fit$lambda)),
     type='n',
     xlab=expression(lambda),
     ylab='train/test error')
points(fit$lambda, err_train_1, pch=19, type='b', col='darkblue')
points(fit$lambda, err_test_1, pch=19, type='b', col='darkred')
legend('topright', c('train','test'), lty=1, pch=19,
       col=c('darkblue','darkred'), bty='n')
colnames(fit$beta) <- paste('lam =', fit$lambda)
print(fit$beta %>% as.matrix)
```

```{r}
## plot path diagram
plot(x=range(fit$lambda),
     y=range(as.matrix(fit$beta)),
     type='n',
     xlab=expression(lambda),
     ylab='Coefficients')
for(i in 1:nrow(fit$beta)) {
  points(x=fit$lambda, y=fit$beta[i,], pch=19, col='#00000055')
  lines(x=fit$lambda, y=fit$beta[i,], col='steelblue')
}
text(x=0, y=fit$beta[,ncol(fit$beta)], 
     labels=rownames(fit$beta),
     xpd=NA, pos=4, srt=45)
abline(h=0, lty=3, lwd=2)
```

